---
editor: 
  markdown: 
    wrap: 72
---

# Data

## Description

In this project, we will use two datasets about electric vehicles in
Washington State. Both datasets are collected and maintained by
Washington State Department of Licensing, and are published on
[Washington Open Data](https://data.wa.gov/). The two datasets are
created on April 16, 2019 and May 8, 2019 respectively and are updated
monthly with the latest update on November 15th.

The first dataset [*Electric Vehicle Population
Data*](https://data.wa.gov/Transportation/Electric-Vehicle-Population-Data/f6w7-q2d2)
contains information about current electric vehicle registrations in WA.
It has 159,467 rows each representing a registered electric vehicle, and
17 columns including VIN (1-10), County, City, State, Postal Code, Model
Year, Make, Model, Electric Vehicle Type, Clean Alternative Fuel Vehicle
(CAFV) Eligibility, Electric Range, Base MSRP, Legislative District, DOL
Vehicle ID, Vehicle Location, Electric Utility, and 2020 Census Tract.
The dataset is clean except for some undocumented entries of Electric
Range and Base MSRP represented as 0, which clearly doesn't make sense.
We will analyze this further in section 2.3.

The second dataset [*Electric Vehicle Population Size
History*](https://data.wa.gov/Transportation/Electric-Vehicle-Population-Size-History/d886-d5q2)
contains monthly updated total count of Plug-In Hybrid Electric Vehicle,
Battery Electric Vehicle, and Electric Vehicle Total in WA from January,
2017 to October, 2023. It has 82 rows and 4 columns. This dataset is
ready-to-use as it has no issues.

Our data source allows us to download csv files which we will then
upload to our github repository for further usage.

## Research Plan

-   Trends in Electric Vehicle Adoption

    -   To identify the patterns of electric vehicle adoption over the
        years, we plan to plot a stacked bar chart based on the time
        series data within [*Electric Vehicle Population Size
        History*](https://data.wa.gov/Transportation/Electric-Vehicle-Population-Size-History/d886-d5q2)
        dataset.
    -   The stacked bar chart would show the size of population owning
        PHEV, BEV, and their sum (EV) overtime.

-   Popular Electric Vehicle Models/Makes

    -   We choose to use the first dataset [*Electric Vehicle Population
        Data*](https://data.wa.gov/Transportation/Electric-Vehicle-Population-Data/f6w7-q2d2),
        which contains detailed information about each registered
        electric vehicle in WA.
    -   We are going to plot three pie charts to show the percentage of
        each make for PHEV, BEV, and EV. A cleveland dot plot is also
        suitable for displaying the ranking of different makes.
    -   We would then draw a faceted bar chart for the count of all
        makes at county level.
    -   After that, we would generate a list of interested makes.
        Therefore, we are able to take a closer look at customer's
        preference by ploting a cleveland dot plot of models for subset
        of makes.

-   Distribution of Electric Range

    -   We would also address this question using [*Electric Vehicle
        Population
        Data*](https://data.wa.gov/Transportation/Electric-Vehicle-Population-Data/f6w7-q2d2).
    -   We intend to create a boxplot for electric range for all EVs. 
    -   A faceted boxplot for different EV type (PHEV & BEV) is also
        necessary to explore the difference.

-   Price Analysis

    -   Similarly, we would use [*Electric Vehicle Population
        Data*](https://data.wa.gov/Transportation/Electric-Vehicle-Population-Data/f6w7-q2d2)to
        investigate this question.
    -   We prepare to draw a boxplot and/or a ridgeline plot for MSRP of
        EV, as well as a faceted plot for PHEV and BEV. 
    -   Moreover, we hope to graph a scatterplot of MSRP vs. electric
        range, separated by EV type. 

## Missing value analysis
```{r echo=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(tibble)
library(redav)
evt <- read.csv("/Users/wy/EV/data/Electric_Vehicle_Population_Size_History.csv")
names(evt) <- c("Date", "PHEV_Count", "BEV_Count", "EV_Count")
plot_missing(evt, percent = TRUE)
```
There's no missing value in *Electric Vehicle Population Size History* dataset, as we can tell from the graph.

```{r echo=FALSE, message=FALSE}
evp <- read.csv("/Users/wy/EV/data/Electric_Vehicle_Population_Data.csv")

evp <- evp %>%
  select(-Legislative.District, -Postal.Code, -`X2020.Census.Tract`, -`VIN..1.10.`,
         -Model.Year, -`Clean.Alternative.Fuel.Vehicle..CAFV..Eligibility`, -DOL.Vehicle.ID, 
         -Vehicle.Location, -Electric.Utility, -State) %>%
  mutate(Base.MSRP = ifelse(Base.MSRP == 0, NA, Base.MSRP), 
         Electric.Range = ifelse(Electric.Range == 0, NA, Electric.Range))

names(evp) <- c("County", "City", "Make", "Model", "Type", "Range", "MSRP")

tidyevp <- evp %>% 
    rownames_to_column("id") %>% 
    gather(key, value, -id) %>% 
    mutate(missing = ifelse(is.na(value), "yes", "no"))

ggplot(tidyevp, aes(x = key, y = id, fill = missing)) +
  geom_tile() + 
  ggtitle("Electric Vehicle Population Missing Values") +
  ylab('') + 
  scale_fill_viridis_d() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r echo=FALSE, message=FALSE}
plot_missing(evp, percent = TRUE)
```
As for *Electric Vehicle Population Data* dataset, it might appear at first that missing values 
only exist in columns that we're not interested in. However, we notice entries with value of "0" for 
Base MSRP and Electric Range, which clearly does NOT make sense and thus should be treated as missing 
values. From the graph we can tell that other variables of interest have no missing values. Almost all values 
of Base MSRP are missing and about a half of Electric Range values are missing, with a extremly low percentage 
of rows not missing values from any columns. We also notice that for some specific model, we have instances for which Base MSRP is missing and instances for which Base MSRP is not missing, similar with Electric Range. We will address 
this issue when preprocessing the data, by filling known missing values using non-missing instances and filling unknown missing values with data provided by official websites of vehicle brands.
